MITM: Logging a Bluetooth Keyboard

### 2013-03-16 final_paper/paper.tex
wrote setup and introduction

\category{H.4}{Bluetooth Attacks}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
\terms{Theory}

@@

\section{Introduction}
Many Bluetooth devices are marketed explicitly for mobile devices such as phones and tablets. 
Attack against tablets not common

threat for MITM attacks is high \cite{sandhya2012analysis}

Many Bluetooth devices are marketed explicitly for mobile devices such as phones and tablets. 
Attack against tablets not common

Public key management is a problem \cite{almonanimitmssp}

@@

\textbf{Security Modes}: Bluetooth devices can be in a variety of different security modes with varying levels of security \cite{ma2007keystroke}. For backwards compatibility reasons the lowest level of security can still be used.

\begin{enumerate}
\item Mode 1: no security or encryption
\item Mode 2: no security until channel is established on L2CAP level (software)
\item Mode 3: has security before link setup on the LMP level (hardware)
\item Mode 4: introduced Secure Simple Pairing, pairing protocol to protect devices while a pairing is completed
\end{enumerate}

@@

\textbf{Numeric Comparison}: a 6 digit number is displayed on both devices and a user confirms that the numbers match (this is designed for devices that both have display and input capabilities)\\\\
\textbf{Passkey Entry}: designed when one device has input (keyboard) and the other mostly display; the device that has the display shows a 6 digit number, which the user types in on the input device.\\\\
\textbf{Just Works}: one device has no input or display; the user must accept the connection without being able to visualize that the numbers are the same (more in nist)\\\\
\textbf{Out of Band (OOB)}: designed to connect when two devices are touched together, (e.g. Near Field Communication or NFC)\\\\
%TODO 5 steps of pairing protocol
\subsection{MITM Bluetooth Attacks}
We focus on man-in-the-middle attacks (MITM) against devices that use Passkey Entry as their primary mode of authentication for pairing. Historically, these attacks have been simple to implement and successful using a variety of approaches.

@@

\subsection{Hardware and Software}
For this attack, we wanted to test several versions of Bluetooth as well as different Operating Systems to determine how effective the attack is against different hardware and software. The following are the devices we used for the attack:

@@

To execute pairing, the Macbook, the Thinkpad Laptop and Tablet all use Passkey Entry authentication where a pin must be entered into the keyboard. Since the attack relies on SSP, which is largely the same across all versions of Bluetooth since it's implementation in v 2.1, the differences in Bluetooth versions across the devices should not matter. The largest variable in whether or not the attack is possible is how the OS handles the link keys and sending multiple link requests from a single address. 
\begin{figure}
\begin{verbatim}
ACL data: handle 11 flags 0x02 dlen 14
L2CAP(d): cid 0x0041 len 10 [psm 0]
A1 01 00 00 04 00 00 00 00 00
\end{verbatim}
\caption{Sample \texttt{hcidump} dump data showing capital A}
\label{fig:mypic}
\end{figure}

\subsection{Setup}
Need a Linux laptop

\subsection{Implementation}
Note about pairing before hand
\begin{verbatim}
sudo hidd --connect <keyboard-address>
\end{verbatim}

\begin{verbatim}
sudo hcidump -x > log.out
\end{verbatim}

\begin{verbatim}
sudo ./hidclient -e8 -x
\end{verbatim}

Connect on the victim pc.
Run the python script.

@@

\section{Mitigation}
We are proposing three countermeasures to prevent against Man in the Middle attacks on Bluetooth keyboards.

\begin{enumerate}
\item Stay out of discoverable mode as much as possible. This way any attacker would need special hardware to perform any kind of fancy attack and take much longer to actually find the keyboards Bluetooth address.
\item Keyboards should require a PIN to pair every time. Keyboards potentially handle much more sensitive data than a device such as a Bluetooth receiver.
\item Bluetooth manager applications should be smarter. If they notice a previously paired keyboard or device with the same name that has a different Bluetooth address that might warrant a red flag to be displayed to the user.
\end{enumerate}

\section{Future Work}
There are several obvious ways to continue the work we completed.

\begin{enumerate}
\item Implement the entire attack. Including the denial-of-service and pairing without the users knowledge.
\item Get Ubertooth and successfully use it to choose victims without having to have prior knowledge about the victim and the keyboard.
\item Implement the entire attack in a single C/C++ application. This would make testing and debugging much easier to perform. Instead of requiring several separate tools combined together.
\item Acquire passwords!
\end{enumerate}

>>>>

\category{H.4}{Bluetooth Attacks}{Security}
\terms{MITM Attack, Bluetooth Security}

@@

\section{Introduction}
Bluetooth technology has proven to be vulnerable since it's adoption in to mainstream devices \cite{scarfone2008guide}, and while the security continues to improve, the threat for Man in the Middle (MITM) attacks against two paired Bluetooth devices remains high \cite{sandhya2012analysis}. In particular, MITM attacks take advantage of the Secure Simple Pairing (SSP) security protocol that was implemented in v2.1 in 2007, which was designed to enhance the security of Bluetooth devices and to simplify the pairing process for end-users. However, our work, and much of the previous work that focused on SSP MITM attacks, strives to show that there are still inherent security flaws in the implementation of the Bluetooth protocol.

Several researchers have focused on what to do with a MITM Bluetooth attacka, particularly when dealing with input devices such as Bluetooth mice \cite{btmice} and Bluetooth keyboards \cite{cuthbertbluetooth,ma2007keystroke}. However, no work has successfully implemented a Bluetooth keylogger that takes advantage of a MITM attack and that does not require physical access to the victim's devices. Our work creates a proof of concept for a Bluetooth keylogger that can be executed using off-the-shelf hardware.

@@

\textbf{Security Modes}: Bluetooth devices can be in a variety of different security modes with varying levels of security \cite{ma2007keystroke}. For backwards compatibility reasons the lowest level of security can still be used.
\begin{enumerate}
\item Mode 1: no security or encryption
\item Mode 2: no security until channel is established on L2CAP level (software)
\item Mode 3: has security before link setup on the LMP level (hardware)
\item Mode 4: introduced Secure Simple Pairing, pairing protocol to protect devices while a pairing is completed
\end{enumerate}

@@

\textbf{Numeric Comparison}: a 6 digit number is displayed on both devices and a user confirms that the numbers match (this is designed for devices that both have display and input capabilities)\\\\
\textbf{Passkey Entry}: designed when one device has input (keyboard) and the other mostly display; the device that has the display shows a 6 digit number, which the user types in on the input device.\\\\
\textbf{Just Works}: one device has no input or display; the user must accept the connection without being able to visualize that the numbers are the same (more in nist)\\\\
\textbf{Out of Band (OOB)}: designed to connect when two devices are touched together, (e.g. Near Field Communication or NFC)\\\\

\subsection{MITM Bluetooth Attacks}
We focus on man-in-the-middle attacks (MITM) against devices that use Passkey Entry as their primary mode of authentication for pairing. Historically, these attacks have been simple to implement and successful using a variety of approaches.

@@

\subsection{Hardware}
For this attack, we wanted to test several versions of Bluetooth as well as different Operating Systems to determine how effective the attack is against different hardware and software. The following are the devices we used for the attack:

@@

To execute pairing, the Macbook, the Thinkpad Laptop and Tablet all use Passkey Entry authentication where a pin must be entered into the keyboard. Since the attack relies on SSP, which is largely the same across all versions of Bluetooth since it's implementation in v 2.1, the differences in Bluetooth versions across the devices should not matter. The largest variable in whether or not the attack is possible is how the OS handles the link keys and sending multiple link requests from a single address. 
\begin{figure}
\begin{verbatim}
ACL data: handle 11 flags 0x02 dlen 14
L2CAP(d): cid 0x0041 len 10 [psm 0]
A1 01 00 00 04 00 00 00 00 00
\end{verbatim}
\caption{Sample \texttt{hcidump} dump data (0x04) showing capital A}
\label{fig:mypic}
\end{figure}

\subsection{Setup}
Our attack setup used two laptops mentioned in the Hardware section and the Linux Command line utilities mentioned in the previous section. We used the Lenovo laptop as the attacker PC and the Macbook Air as the victim. Since the attack is OS ambivalent, the attack could theoretically be executed on a variety of Bluetooth enabled devices. However, given out limited resources, we were only able to attack the Macbook Air and another Desktop running Ubuntu 12.04 successfully.

\subsection{Implementation}
Our worked under the assumption that the pairing process was performed successfully (as prior work suggests is possible) and the attacker is ready to connect and log the keystroke data. To being the keylogging attack, the attacker first scans for the Bluetooth address of both the victim PC and the keyboard by using \texttt{hcitool scan}. Once both the address are located, the attacker pairs with both and connects to the devices to begin data transfer.

To connect to the keyboard in unencrypted mode, the attacker uses the following command, where the <keyboard-address> is the Bluetooth address of the victim keyboard:

\texttt{sudo hidd --connect <keyboard-address>}

Once the attacker has connected to the keyboard, he or she must run the \texttt{hcidump} in another shell and dumps the output into a log file, using the following command:

\texttt{sudo hcidump -x > <log-file>}

The \texttt{-x} option dumps the data in hexidecimal encoding, which is easily extracted later.

The attacker then runs the \texttt{hidclient} with the \texttt{-l} option, which lists the available inputs that can be forwarded to the victim PC.

\texttt{sudo hidclient -l}

The Bluetooth keyboard appears as one of the options, and the attacker must simply replace the X in the following command to specify the input source as the keyboard.

\texttt{sudo ./hidclient -eX -x}

At this point, the victim connects to the attacker as keyboard input, completing the connection. The \texttt{hidclient} forwards the data from the keyboard which is being logged by the \texttt{hcidump} on the attacker's machine. As mentioned before, this attack relies on the victim being tricked into pairing and connecting to the attacker's PC. For the purposes of this attack, we connected from the victim's PC using the \texttt{hidd} command in Linux. In a real-world attack, the victim should be able to connect to the attacker "keyboard" using the OS's built-in device manager.

The attack continues until the attacker or victim severs the connection, at which point the attacker runs a simple script to extract the data. We used a simple Python script to decode the output with excellent success.

@@

\section{Mitigation}
We are proposing three countermeasures to prevent against Man in the Middle attacks on Bluetooth keyboards.
\begin{enumerate}
\item Stay out of discoverable mode as much as possible. This way any attacker would need special hardware to perform any kind of fancy attack and take much longer to actually find the keyboards Bluetooth address.
\item Keyboards should require a PIN to pair every time. Keyboards potentially handle much more sensitive data than a device such as a Bluetooth receiver.
\item Bluetooth manager applications should be smarter. If they notice a previously paired keyboard or device with the same name that has a different Bluetooth address that might warrant a red flag to be displayed to the user.
\end{enumerate}

\section{Future Work}
There are several obvious ways to continue the work we completed. 
\begin{enumerate}
\item Implement the entire attack. Including the denial-of-service and pairing without the users knowledge.
\item Get Ubertooth and successfully use it to choose victims without having to have prior knowledge about the victim and the keyboard.
\item Implement the entire attack in a single C/C++ application. This would make testing and debugging much easier to perform. Instead of requiring several separate tools combined together.
\item Acquire passwords!
\end{enumerate}
